---
title: "R Bridge Final"
author: "Ken Howes"
date: "August 6, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## install.packages("hflights")
## This dataset contains all flights departing from Houston airports IAH (George Bush Intercontinental) and HOU (Houston Hobby).


```{r}
library(hflights)
library(ggplot2)
library(gcookbook)
```

## Explore the Dataset
```{r}
dim(hflights)
str(hflights)
head(hflights)
summary(hflights)

table(hflights$UniqueCarrier)
```

## Create a subset of columns and explore data further

```{r}
hflights2 <- subset(hflights, ,select=c(UniqueCarrier, ArrDelay, DepDelay))

dim(hflights2)
str(hflights2)

attach(hflights2)
par(mfcol=c(1,2))
boxplot(hflights2$ArrDelay)
boxplot(hflights2$DepDelay)
```

## Create a vector with the average arrival delay by carrier
```{r}
ArrDelayMeans <- aggregate(ArrDelay ~ UniqueCarrier, hflights, mean)
head(ArrDelayMeans)
```

## Which carriers have the best on-time performance (recipe from R Graphics Cookbook)?
```{r}
ggplot(ArrDelayMeans, aes(x=ArrDelay, y = UniqueCarrier)) + geom_point()          # original attempt

ggplot(ArrDelayMeans, aes(x=ArrDelay, y = reorder(UniqueCarrier, ArrDelay))) +    # order by average delay
  geom_point(size = 3) +                                                          # use a larger dot
  theme_bw() +                                                                    # streamline look of grid
  theme(panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(colour="grey60", linetype="dashed")) 
```
